\chapter*{Abstract}
Steadily increasing use of computers, programs, websites and mobile applications creates~a~demand for tools and solutions that monitor their operation and analyze user behavior. Those tools support developers in improving the quality of their products and help customers achieve their goals. At the same time~a~new category of products that interprets and uses the sensor signals is starting~to~emerge. This is largely due~to~the recent technological advancements, increase in the available computing power and the falling prices of electronic components.

This thesis summarizes the current state of development of applications and frameworks that monitor devices, programs and user behavior. Examples of solutions from both academia and the commercial sector are discussed and compared. Their diversity is visible not only in terms of their purpose, but also construction and operation. Systematic literature review of scientific publications is conducted~to~assess the progress in research~on~the use of sensors~to~detect human activities. Except for~a~handful of original publications, most of the analyzed research papers approach~a~common problem of detecting the same set of basic activities in attempt~to~achieve an increase in accuracy.

The second part of this thesis focuses~on~the technique of creating heat maps from user interactions with mobile applications. Creation process of an original heat map tool is described in detail, starting with the design and architecture, going through the implementation and encountered challenges and finishing with tests and release. The final part of this thesis describes the evaluation process of the created solution. A~mobile application extended with the proposed tool was distributed~to~a group of testers. Collected data was processed into heat maps which were then used~to~identify issues with the application user interface, come up with good design practices and better understand other potential uses of such tool. Visualizing interactions using heat maps provides~a~unique insight into the application~as~it is seen through the eyes of users and allows~to~draw important conclusions about its interface. \\

\noindent\textbf{Keywords:} monitoring, activities, analysis, users, behavior, sensors, applications, heat maps, interactions, visualization, interface, accessibility, processing, evaluation
